A Sound and Complete Deep Embedding of Arbitrary Normal Modal Logics in Coq

### 2025-08-20 paper/COLA/paper.tex
Change some snippets

In the folllowing, we omit the $\Gamma$ subscripts. The canonical model is formalized in Coq as follows: first, we define the type \mintinline{coq}{W} that pack
the construction of the maximal consistent sets for a theory \mintinline{coq}{G}.
\begin{minted}{coq}
Inductive W: Type :=
| W_mk D: Consistent A D -> Maximal D -> Subset G D -> W.
\end{minted}
Next, we encode the accessibility relation, $\mathcal{R}$, by
the definition \mintinline{coq}{canonical_accessibility} and use
notation \mintinline{coq}{R} for it. Observe that it is just the
translation of the mathematical notation into Gallina's code.
\begin{minted}{coq}
Definition canonical_accessibility: relation W :=
  fun w v => forall p, w [! []p !] -> v p.

@@

\paragraph{The completeness proof} We almost have all the pieces to prove the completeness theorem. The next lemma
shows that a formula holds in the canonical model if, and only if, it is provable in the deductive system.
\begin{minted}{coq}
Lemma determination: forall p, (M |= p) <-> (A; G |-- p).
\end{minted}
The proof is divided in two more lemmas for each implication direction using the truth lemma. Finally, the desired completeness theorem is stated simply as follows:
\begin{minted}{coq}
Theorem completeness: forall p, (G ||= p) -> (A; G |-- p).
Proof.
  intros p.
  apply contrapositive; intros.
  - apply classic.
  - assert (Consistent A G).
    + apply nonderivation_implies_consistency with p.
      assumption.
    + assert ((M |= p) -> False); intros.
      * apply determination with p in H2; auto.
      * apply H2, H0.
        intros q ? w; apply truth.
        destruct w as (w, ?H, ?H, ?H); simpl.
        apply H6; assumption.
Qed.
\end{minted}
The proof proceeds by the contrapositive.
Suppose that \mintinline{coq}{A; G |-- p} does not hold.
As this system can't derive every formula, it has to be consistent, which we can safely assume.
Since \verb|G| is consistent, it
can be extended into the canonical model \verb|M|. By the fact that
\mintinline{coq}{A; G |-- p} is false, by the determination lemma, we know
that \verb|p| is not true on the canonical model \verb|M|. Conclusion follows
by the truth lemma.

>>>>

In the folllowing, we omit the $\Gamma$ subscripts. The canonical model is formalized in Coq as follows: first, we define the type \mintinline{coq}{W} that pack
the construction of the maximal consistent sets for a theory.
\begin{minted}{coq}
Inductive W: Type :=
| W_mk D: Consistent A D -> Maximal D -> W.
\end{minted}
Next, we encode the accessibility relation, $\mathcal{R}$, by
the definition \mintinline{coq}{canonical_accessibility} and use
notation \mintinline{coq}{R} for it. Observe that it is just the
translation of the mathematical notation into Gallina's code.
\begin{minted}{coq}
Definition canonical_accessibility: relation W :=
  fun w v => forall p, w [! []p !] -> v p.

@@

\paragraph{The completeness proof} We almost have all the pieces to prove the completeness theorem. The next lemma
shows that a formula holds in the canonical model if, and only if, it is provable in the deductive system.
\begin{minted}{coq}
Lemma determination: forall p, (M |= p) <-> (A; Empty |-- p).
\end{minted}
The proof is divided in two more lemmas for each implication direction using the truth lemma. Finally, the desired completeness theorem is stated simply as follows:
\begin{minted}{coq}
Theorem completeness: forall p, (G ||= p) -> (A; G |-- p).
Proof.
  intros p.
  apply contrapositive; intros.
  - apply classic.
  - assert (Consistent A G).
    + apply nonderivation_implies_consistency with p.
      assumption.
    + assert ((M |= p) -> False); intros.
      * apply determination with p in H2; auto.
      * apply H2, H0.
        intros q ? w; apply truth.
        destruct w as (w, ?H, ?H, ?H); simpl.
        apply H6; assumption.
Qed.
\end{minted}
The proof proceeds by the contrapositive.
Suppose that \mintinline{coq}{A; G |-- p} does not hold.
As this system can't derive every formula, it has to be consistent, which we can safely assume.
Since \verb|G| is consistent, it
can be extended into the canonical model \verb|M|. By the fact that
\mintinline{coq}{A; G |-- p} is false, by the determination lemma, we know
that \verb|p| is not true on the canonical model \verb|M|. Conclusion follows
by the truth lemma.
